---
title: "hw01_junco"
author: "Zhander Junco"
format: html
execute:
  echo: true
  warning: false
  message: false
---

## Instructions

1.  **Clone the Homework Repository**\
    Clone the GitHub repository containing the homework file to your **local machine**.

2.  **Rename the File**\
    Rename the `.qmd` file from `hw01.qmd` to: `hw01_yourusername.qmd`.\

For example: `hw01_semiyari.qmd`

### Working on Your Homework

-   As you complete each part of the homework:

    -   **Stage** and **commit** your `.qmd` file *after each question/sub-question*, such as `1.1`, `1.2`, etc.\
    -   **Render** your `.qmd` file to **HTML only** (no PDF or Word formats).\
    -   **Push** your updates to GitHub after completing each **main question**, such as `1.0`, `2.0`, etc.

> Committing and pushing to GitHub **are graded components** of this assignment.

## Final Checks Before Submission

-   **Re-render** to HTML before your final commit and push.\

-   **Check your output:**\

    -   Make sure **all images display correctly**.\
    -   Avoid including **excessive output or large datasets** unless asked.

## Final Submission

Once you have finished and pushed your final version:

-   Copy the link of your **GitHub Pages site** (where the rendered HTML is published).
-   **Post the link** to the Canvas assignment submission page.

```{r}

library(pacman)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(dplyr)

pacman::p_load(pacman, tidyverse,
               nycflights13)
```

### Consider the data in the nycflights13 package.

```{r}
flights <- flights
```

### **Problem 2**: Is there a relationship between the age of a plane and its delays?

```{r}

planes_age <- planes %>%
  mutate(age = 2013 - year)

```

```{r}
flights_age <- flights %>%
  inner_join(planes_age, by = "tailnum") %>%
  mutate(dep_delay = ifelse(dep_delay < 0,0, dep_delay))

age_delay <- flights_age %>%
  group_by(age) %>%
  summarise(
    mean_delay = mean(dep_delay, na.rm = TRUE),
    median_delay = median(dep_delay, na.rm = TRUE),
    n = n ()
    )


```

```{r}

age_delay %>%
  ggplot(aes(age, mean_delay)) +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    x = "Age of Aircraft (years)",
    y = "Mean Departure delay (minutes)", 
    title = "Older planes show only a weak uptick in average delay"
  )
  

```

Plane age is not a reliable predictor of departure delay; if anything, newer aircraft experience slightly higher average delays than old ones, but the effect is small relative to the overall variability. i.e,, delays peak (17-18 min) around 10-15 years, then drift downward for aircraft older than \~30.

### **Problem 3**: Find the 10 days of the year that have the highest median departure delay, then select all flights from those 10 days.

```{r}
worst_day<- flights%>%
  group_by(month, day)%>%
  summarise(median_dep = median(dep_delay, na.rm = TRUE),
            n = n(),
            .groups = "drop") %>%
  slice_max(order_by = median_dep, n = 10, with_ties = FALSE)%>%
  arrange(desc(median_dep))

worst_day
```

```{r}
flights_worst <- flights%>%
  semi_join(worst_day, by = c("month", "day"))%>%
  count(month, day, sort = TRUE)

flights_worst

```
