---
title: "Functions-Iterations-Dates-purr"
format: html
editor: visual
---

## Instructions

Complete the following exercises using the `lubridate` and `purrr` packages in R. Ensure that your solutions are optimized and use functional programming principles where applicable.

1.  Load the necessary libraries.
2.  Answer each question in separate R code chunks.
3.  Provide detailed explanations for your approach.
4.  Submit the rendered HTML file.

```{r}
library(lubridate)
library(purrr)
```

## Exercise Functions:

### Question 1

Write a function (using r code and structure demonstrated in class) to calculate a z score for a given observed value, a mean, and a standard deviation value. And then use your function to find a z score for the following problem. (Research the internet to find the formula used to calculate a z score)\
Observed value = 25.77, mean = 23.54, standard deviation = 2.442

#### Z-score formula: 

$$ z= \frac{(x-\mu)}{\sigma}$$

```{r}
# Function to calculate z-score 

calculate_z <-function(observed, mean, sd) {
  z <-(observed-mean)/sd
  return(z)
}

```

```{r}
# Given values 
observed_value <- 25.77
mean_value <- 23.54
sd_value <- 2.442

# Calculate z-score

z_score_result <- calculate_z(observed_value,mean_value,sd_value)
z_score_result
```

The observation is \~ 0.91 standard deviations above the mean.

### Question 2

Write a function to calculate the natural log of a number multiplied by the common log of the same number divided by the cube root of a given prime number, $\dfrac{\ln x \times \log x }{\sqrt[3]{p}}$.\
Use your function to find the answer if the number to be used for both log expressions is 32 and the given prime number is 11, $x = 32, \;\;p=11$.\
Also use R code to round your answer to the nearest tenth

```{r}
# Function to calculate natural log of a number 

calculate_expression <- function(x,p) {
  numerator <- log(x) * log10(x)
  denominator <- p^(1/3)
  result <- numerator/denominator
  rounded_result <- round(result, 1)
  return(rounded_result)
}
  
```

```{r}
# Function applied to the answer 

calculate_expression(32,11)

```

By defining the function that combines both natural and common logarithmic operations, I was able to compute the expression without repeating steps or hardcoding values. The use of log () for the natural log, log10() for the common log, and exponentiation for the cube root kept the logic straightforward and readable. \
\
By using function, I've made it easy to reuse for other values of x and p. The final result, rounded to the nearest tenth, was 2.3.

### Question 3

Use and show R coding to calculate the standard deviation for each variable of the data table `mtcars` using the `Special For Loop Method`, which includes three components:

-   Output – an object to store the results

-   Sequence – the set of items to loop through

-   Body – the operations performed in each iteration

```{r}
sd_vec <- vector(mode = "numeric", length = length(mtcars))
names(sd_vec) <- colnames(mtcars)
for (i in seq_along(mtcars)) { 
  sd_vec [[i]] <- sd(mtcars[[i]], na.rm = TRUE)
} 

sd_vec
```

### Question 4:

Write only one line of the code to return only the leap years fo years 2024, 3021, 2000 or 1800, 800, 5050, 600 and 7156. using subsetting with `[]` and without using built-in functions.

A leap year is divisible by 4, except years divisible by 100 unless also divisible by 400. So the condition is:

```{r}
# Define years: 
years <- c(2024, 3021, 2000, 1800, 800, 5050, 600, 7156)
```

```{r}
# Code:
years[years %% 4 == 0 & (years %% 100 != 0 | years %% 400 == 0)]
```

To solve this problem, I first defined the vector of years in one chunk. Then, I used a one line expression with logical subsetting to filter only the years that meet the leap rule: divisible by 4, except when divisible by 100, unless also divisible by 400. This approach avoids any built in helper functions and clearly expresses the logic required for leap year classification.

## Advanced Date Manipulation with lubridate

### Question 5:

Generate a sequence of dates from **January 1, 2015** to **December 31, 2025**, spaced by every **two months**. Extract the **year**, **quarter**, and **ISO week number** for each date.

```{r}
# Step 1: Date sequence

date_seq <- seq(from = ymd("2015-01-01"), to = ymd("2025-12-31"), by = "2 months")
```

```{r}
# Step 2: Year, quarter and ISO week number

date_info <- data.frame(
  date = date_seq,
  year = year(date_seq),
  quarter = quarter(date_seq),
  iso_week = isoweek(date_seq)
)

date_info

```

In this question, I used seq() and lubridate to generate a sequence of dates starting from January 1, 2015, to December 31, 2025, with spacing of every two months. I then extracted useful time related components like year, quarter , and ISO week number using year(), quarter () , and isoweek(). This kind of transformation is particularly helpful in time series or panel data, where grouping and aggregating by these calendar units can support better insights.

### Question 6:

Given the following dates, compute the difference in **months** and **weeks** between each consecutive pair.

``` r
sample_dates <- c("2018-03-15", "2020-07-20", "2023-01-10", "2025-09-05")
```

```{r}
# Vector dates defined 
sample_dates <- ymd(c("2018-03-15", "2020-07-20", "2023-01-10", "2025-09-05"))

```

```{r}
# Difference in time between each pair
date_diffs <- diff(sample_dates)

# Difference in weeks

week_diffs <- as.numeric(date_diffs, units = "weeks")

# Difference in months 

month_diffs <- time_length(date_diffs, unit = "months")
```

```{r}

# Data frame 

date_diff_df <- data.frame(
  from = sample_dates [-length(sample_dates)],
  to = sample_dates [-1],
  weeks = round(week_diffs, 1),
  months = round(month_diffs, 1)
)

date_diff_df
```

Here, I calculated the difference between each consecutive pair of dates from a short list. Using diff () on the date vector gave me interval objects, which I then converted to weeks ( with as.numeric ) and months (using time_length). This approach allowed me to clearly compare time gaps between events in multiple units. It's a practical use case for handling time intervals in reporting or time to event analysis.

### Question 7: Higher-Order Functions with purrr

Using `map()` and `map_dbl()`, compute the **mean**, **median**, and **standard deviation** for each numeric vector in the following list:

``` r
num_lists <- list(c(4, 16, 25, 36, 49), c(2.3, 5.7, 8.1, 11.4), c(10, 20, 30, 40, 50))
```

```{r}
num_lists <- list(c(4, 16, 25, 36, 49), c(2.3, 5.7, 8.1, 11.4), c(10, 20, 30, 40, 50))
```

```{r}

# Mean for each vector

means <- map_dbl(num_lists, mean)

# Median for each vector 

medians <- map_dbl(num_lists, median)

# Standard Deviation for each vector

sds <- map_dbl(num_lists, sd)
```

```{r}
# Summary Data frame 

summary_df <- data.frame(
  set = paste0("Vector_", seq_along(num_lists)),
  mean = round(means, 2),
  median = round(medians, 2),
  sd = round(sds, 2)
)

summary_df
```

I used map_dbl () to apply statistical summary functions across each numeric vector in the list. This approach avoids the need for loops and keeps the code clean and readable. It's specially useful for quickly summarizing structured data. \

### Question 8: Combining lubridate and purrr

Given a list of mixed date formats, use `map()` and `possibly()` from `purrr` to safely convert them to **Date** format and extract the **month name**.

``` r
date_strings <- list("2023-06-10", "2022/12/25", "15-Aug-2021", "InvalidDate")
```

```{r}
date_strings <- list("2023-06-10", "2022/12/25", "15-Aug-2021", "InvalidDate")

safe_parse <- possibly(
  ~ parse_date_time(.x, orders = c("ymd", "Ymd", "d-b-Y", "Y/m/d")),
  otherwise = NA
)

parsed_dates <- map(date_strings, safe_parse)
parsed_vec <- as_date(unlist(parsed_dates))

month_names <- map_chr(parsed_dates, function(x) {
  if (is.na(x) || length(x) == 0) NA_character_
  else as.character(month(x, label = TRUE, abbr = FALSE))
})

date_df <- data.frame(
  original = unlist(date_strings),
  parsed_date = parsed_vec,
  month_name = month_names,
  stringsAsFactors = FALSE
)

date_df
```

In this task I worked with a list of date strings written in different formats. To handle this safely, I used purrr along with possibly to attempt to parse each date while handling failures by returning NA. I used parse_date_time to support multiple input formats. After parsing the valid dates, I extracted the full month name using month and wrapped it in as.character, to avoid warnings. This approach keeps the pipeline robust and flexible, even when dealing with messy or inconsistent data, which is often the case in real world datasets. \

## Submission

-   Save your Quarto file after completing each problem, render it, then stage and commit the changes.\

-   Include well-commented code and clear explanations.\

-   Make at least **one commit per question**.\

-   Submit the rendered file to **Canvas** and push it to your **GitHub repository**.\

-   Provide the link to your GitHub repository in the **Canvas gradebook** as a text submission.\

🔹 **Each problem is worth 10 points.**\

🔹 You will receive an additional **20 points** if you correctly push your completed file to GitHub and make at least one commit for each question.\

**Good luck!**
