---
title: "Homework Assignment: Sentiment Analysis of Emma"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.align  = "center",
                      fig.height = 3, fig.width = 4)
ggplot2::theme_set(ggplot2::theme_bw() + ggplot2::theme(strip.background = ggplot2::element_rect(fill = "white")))
```

# Exploring Joyful Language in Jane Austen’s *Emma* using Tidytext

### **Instructions**

Complete the following exercises using the appropriate packages in R. Ensure that your solutions are optimized and use functional programming principles where applicable.

1.  Load the necessary libraries.
2.  Answer each question in separate R code chunks.
3.  Provide detailed explanations for your approach.
4.  Submit the rendered HTML file.

#### **Objective**:

Use the **`tidytext`** package and three different **sentiment lexicons** (`nrc`, `afinn`, `bing`) to explore **positive/joyful words** in *Emma* by Jane Austen. You will tokenize the text, apply sentiment filters, visualize frequent sentiment words using `ggplot2`, and create a word cloud.

### **Tasks**:

1.  **Data Preparation**

a\. Load the `austen_books()` dataset from the **`janeaustenr`** package.

```{r}
library(janeaustenr)
library(tidyverse)
library(dplyr)
library(stringr)
library(tibble)
library(tidyr)
library(tidytext)
library(textdata)
```

b\. Group by book and detect chapter boundaries using regex; and Create `linenumber` and `chapter` columns.

```{r}

austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(
      text, 
      regex("^chapter\\s+[\\divxlc]+", 
            ignore_case = TRUE
           )
      )),
    .before = text
  ) %>%
  ungroup() %>%
  select(book, chapter, linenumber, text) -> 
  chapter_boundaries
```

***The code processes Jane Austen's text by:***

-   Grouping the data by book title

-   Creating two new columns:

    -   Line number: assigns a sequential number to each line

    -   chapter: detects chapter headers using regex and assigns a chapter number using cumulative counting with cumsum()

-   The regex looks for lines that begin with the word "chapter" followed by a space and a Roman numeral or digit.

-   These new columns are inserted before the text column using .before = text.

-   The grouping is removed (ungroup()), and only relevant columns (book, chapter, linenumber, text) atr selected.

-   The final processed data is saved into an object called chapter_boundaries

***Regex Breakdown:***

-   \^chapter: matches lines that start with the word "chapter"

-   \\\\s+: matches one or more spaces after "chapter"

-   \[\\\\divxlc\] +: matches Roman numerals, handling lowercase, and could also allow digits depending on variation. It also assigns them as chapter boundaries.

Chapter boundaries refer to the starting points of new chapters (we're not trying to figure out where they end).

2.  **Tokenization**

-   Use `unnest_tokens()` to tokenize text into individual words.
-   Explain briefly **why we name the output column `word`** (include this as a comment in your script).

```{r}
## We name the output column 'word' because is the conventional tidytext function for sentiment analysis

chapter_boundaries %>% 
  unnest_tokens(word, text) %>%  ## Tokenize text into individual words
  mutate(word = str_extract(word, "[a-z']+")) %>%
  anti_join(stop_words, by = "word") -> 
  tidy_author 
```

***Code Explanation:***

-   We name the output column "word" in unnest_tokens(word,text) because that's the standard name that functions in tidytext. It makes later joins with sentiment lexicons and stop word list easier, since those also use "word" as the column name.

-   We use str_extract() to keep only alphabetic characters and apostrophes, removing punctuation or formating symbols. Then we remove common stop words (like "the", "and", "of") using anti_join(stop_words), which helps focus the analysis on meaningful words.

3.  **Sentiment Analysis**

-   Filter joy/positive words from **each** of the three sentiment lexicons:

    -   `nrc` (joy)

    ```{r}
    nrc_joy <- get_sentiments("nrc") %>% 
      filter(sentiment == "joy")
    ```

    -   `afinn` (positive scores ≥ 1)

    ```{r}
    afinn_positive <- get_sentiments("afinn") %>% 
      filter(value >= 1)
    ```

    -   `bing` (positive)

    ```{r}
    bing_positive<- get_sentiments("bing") %>% 
      filter(sentiment == "positive")

    ```

-   Join each with *Emma*'s text and:

    -   Count word frequency.
    -   Filter for frequently occurring words (`n > 50`).
    -   Visualize using a **bar chart** (`ggplot2`) and a **word cloud** (`wordcloud`).

1.  **Push to GitHub**

    -   Push your complete R script (`.R` file) to a **GitHub repository**.

    -   Your script should include:

        -   Data wrangling
        -   Sentiment filtering and joins
        -   Visualization code

    -   You must include **at least 5 meaningful commits** to document your workflow.

------------------------------------------------------------------------

## **Rubric 100 Points**

| Category                                       | Points  |
|------------------------------------------------|---------|
| Step 1: Data wrangling (linenumber, chapter)   | 10      |
| Step 2: Tokenization + explanation             | 10      |
| Step 3a: NRC sentiment analysis + plots        | 10      |
| Step 3b: AFINN sentiment analysis + plots      | 10      |
| Step 3c: BING sentiment analysis + plots       | 10      |
| Code readability, structure, and comments      | 10      |
| Word cloud for each lexicon                    | 20      |
| GitHub submission                              |         |
| Commit history (at least 5 meaningful commits) | 20      |
| **Total**                                      | **100** |

Good Luck!
